A Method for the Spectroscopic Inference of Stellar Parameters

### 2014-07-17 tex/paper.tex

\newcommand{\vM}{\mathsf{M}}
\newcommand{\vD}{\mathsf{D}}
\newcommand{\vR}{\mathsf{R}}
\newcommand{\vC}{\mathsf{C}}
\newcommand{\fM}{ \vec{{\bm M}}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ \vec{{\bm D}}}
\newcommand{\fDi}{D_i}
\newcommand{\fR}{ {\bm R}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}
\newcommand{\matern}{Mat\'{e}rn}

@@

There are various approaches to synthesizing a spectrum, $\flam$, for a specific set of model
parameters, $\fundamentalParameters$.  In an ideal case, a model stellar atmosphere is constructed and then 
subsequently processed through a radiative transfer code \citep[e.g.,][]{kurucz93,hauschildt99}.  
However, in general this approach is still computationally prohibitive for any iterative method of 
probabilistic inference.  One partial compromise is to interpolate over a library of atmosphere 
structures that were pre-computed for a discrete grid of parameter values, $\gridParameters$, for some arbitrary $\fundamentalParameters$, and then perform a radiative transfer calculation with 
that interpolated atmosphere to synthesize $\flam$ \citep[e.g., as for SME;][]{valenti96}.  A more 
common variant is to instead rely on interpolation over a library of pre-synthesized model spectra, 
$\flam(\gridParameters)$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While 
technically the former approach is most similar to the ideal case, the computational cost of 
repeated spectral synthesis is sufficiently high to make a detailed exploration of parameter space 
(particularly for data with a large spectral range) considerably less appealing.  A related, but 
different approach is to eschew forward modeling entirely (and therefore repeated spectral 
syntheses and/or library interpolations), and instead evaluate the models only at the discrete grid 
points of the library.  Then, these discretized samples of the posterior probability density can be 
interpolated to an arbitrary $\fundamentalParameters$ to construct appropriate confidence intervals 
\citep[e.g., the method of SPC;][]{buchhave12}.  The difficulty with this latter approach is that the 
parameter uncertainties can be smaller than the grid spacing; in that case, there is valid concern 
that this interpolation might not accurately recover intrinsic parameter degeneracies.

Here we opt to take the computationally expedient approach that employs a library of model spectra, 
$\flam(\gridParameters)$, where $\fundamentalParameters = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
(in practice, the metallicity $Z$ is often parameterized by the iron abundance, [Fe/H]).  However, 
it is worth noting that the techniques we will develop are applicable to {\it any} approach for 
generating a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
$\vt_{\ast}$ must be interpolated from among the spectral library, 
\begin{equation}
\flam(\gridParameters) \leadsto \flam(\fundamentalParameters),
\label{eqn:interp}
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesireable level of inaccuracy (particularly in the $Z$ dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
calculated location in $\gridParameters$, and then comparing the interpolated spectrum 
with the corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\gridParameters$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
be accurate within a few percent per high resolution model pixel.  Ideally, this pre-interpolation 
could be avoided if the spectral library was computed over a refined grid (with a substantial 
up-front computational investment); but for the time being, we can empirically propagate these 
interpolation uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).

@@

We can further divide $\observationalParameters$ into those parameters that impact the model primarily in the 
spectral or flux dimensions; $\observationalParameters = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
former, we consider three kernels that contribute to the observed line-of-sight velocity 
distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, treats the instrumental 
spectral broadening.  For illustrative purposes we assume $\mathcal{F}_v^{\rm inst}$ is a Gaussian 
with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm rot}$, characterizes the 
broadening induced by (projected) stellar rotation, parameterized by $v\sin{i}$ as described by 
\citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm dop} = \delta(v-v_z)$, 
incorporates the radial velocity through a Doppler shift.  The model spectrum is modified by the 
parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
a convolution in velocity-space,
\begin{equation}
\flam(\fundamentalParameters, \vt_{{\rm obs}, v}) = \flam(\fundamentalParameters) \otimes \varphi_v = \flam(\fundamentalParameters) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
\label{eqn:broadening}
\end{equation}
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation}
\flam(\fundamentalParameters, \vt_{{\rm obs}, v}) \mapsto \vM(\fundamentalParameters, \vt_{{\rm obs}, v}),
\label{eqn:resampling}
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-dimensional array $\vM$ (and $N_{\rm pix}$ is the number of pixels in the spectrum).  
For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
post-processing steps.

\begin{figure}[!htb]
\begin{center}
  \includegraphics{figs/kernels.pdf}
  \includegraphics{figs/high2low.pdf}
  \caption{Panel 1: the instrumental (${\rm FWHM} =6.8\;\kms$) and rotational ($v \sin i = 5\;\kms$) broadening kernels, and their composite. Panel 2: A section of the raw, high resolution synthetic model at $R=$500,000. Panel 3: The spectrum after applying the composite broadening kernel, $\varphi_v$. Panel 4: The broadened spectrum downsampled to the pixels of the detector.}
\label{fig:broadening}
\end{center}
\end{figure}

At this stage, the model is further modified in the flux dimension.  A typical synthetic spectrum 
is computed as the flux that would be measured {\it at the stellar surface}, and so needs to be 
diluted by the subtended solid angle, $\Omega = (R_{\ast}/d)^2$, where $R_{\ast}$ is the stellar 
radius and $d$ is the distance.  An additional wavelength-dependent scaling factor is applied to 
account for interstellar extinction, assuming a 
derived extinction law $A_{\lambda}$ \citep[e.g.,][]{cardelli89} that is parameterized by $A_V$.  
The parameters $\vt_{{\rm obs}, f} = [\Omega, \,\, A_V]$ are applied as
\begin{equation}
\vM(\vt) \equiv \vM(\vt_{\ast}, \vt_{\rm obs}) = \vM(\vt_{\ast}, \vt_{{\rm obs}, v}) \times \Omega \times 10^{-0.4\,A_{\lambda}},
\label{eqn:scaling}
\end{equation}
where we have simplified the notation by writing $\vt \equiv [\vt_{\ast}, \,\, \vt_{\rm obs}]$.

@@

We adopt a somewhat analogous approach to deal with this issue, but cast it internal to the 
modeling framework to appropriately propagate the uncertainty introduced by additional degrees of 
freedom in the model.  The residual calibration errors are treated as an explicit contribution to 
the model spectrum, enabling us to explore the distribution of possible calibrations (see 
Section~\ref{subsec:MCMC}) and then eventually marginalize out the associated nuisance parameters.  
In essence, the inferences for the relevant stellar parameters will properly account for the 
underlying uncertainty in the calibration process.  For a continuous spectrum,
such as a long-slit spectrum or a single spectral order of an echelle spectrum,
this is achieved in practice by distorting the model spectrum with a
(low-degree) Chebyshev polynomial \citep[e.g.,][]{eisenstein06,koleva09}, 
\begin{equation}
\vM(\stellarParameters, \cheb) = \vM(\stellarParameters) \times \sum_n c_n \, T_n,
\label{eqn:chebyshev}
\end{equation}
where $T_n$ are the standard Chebyshev functions of degree $n$ and the coefficients are treated as a
set of nuisance (hyper-)parameters, $\cheb = \{c_0, c_1, \ldots, c_{n-1} \}$. For an echelle spectrum with $N$ spectral orders, there is a Chebsyhev polynomial for each order of the spectrum $i$, such that the total collection of Chebyshev parameters is $\Cheb = \{\chebi{1}, \chebi{2}, \ldots, \chebi{N} \}$.  With judicious priors on $\Cheb$, we can ensure that the 
unintended treatment of real spectral features (e.g., broad, deep molecular bands) as calibration 
artifacts is negligible (see Section~\ref{sec:examples} for examples).  The lowest-degree (scaling) 
coefficient, $c_0$, is by its nature degenerate with the solid angle parameter, $\Omega$.  
Therefore, we enforce an additional constraint by requiring that the mean of the polynomial is 
unity.  For data with a single spectral order, this means setting $c_0 = 1$.  If the goal is to 
model multiple spectral orders, we assign $c_0 = 1$ in an arbitrary order as an anchor, but permit 
the $c_0$ values in other orders to be different (as necessary).  It is worth noting that this 
formalism can, in principle, be extended to develop models for completely uncalibrated spectra, 
rather than the residual calibration mismatches as described here (with a suitable relaxation of 
the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:chebyshev} offers a practical demonstration of how 
these nuisance parameters are applied.

@@

The quality of the model spectrum is assessed by comparing to the data with a pixel-by-pixel 
likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding 
$N_{\rm pix}$-dimensional residual spectrum can be defined for any input parameter set,
\begin{equation}
\vR \equiv \vR(\stellarParameters, \cheb) \equiv \vD-\vM(\stellarParameters, \cheb).
\end{equation}
To quantify the probability of the data conditioned on the model, we adopt a standard 
multi-dimensional Gaussian likelihood function,
\begin{equation}
p(\vD|\vM) =  \frac{1}{[(2 \pi)^{N_{\rm pix}} \det(\vC)]^{1/2}} \exp\left ( -\frac{1}{2}
   \vR^\trans \vC^{-1} \vR \right ), 
   \label{eqn:likelihood}
\end{equation}
which penalizes models that yield larger residuals and explicitly allows for covariances in the 
residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix, $\vC$.  For practical 
numerical reasons, we generally consider the log-likelihood as the fit quality metric, which in 
this case is
\begin{equation}
  \ln{p(\vD | \vM)} = -\frac{1}{2} \left( \vR^\trans \vC^{-1} \vR + \ln{\det{\vC}} + N_{\rm pix} \ln{2\pi} \right).
  \label{eqn:lnlikelihood}
\end{equation}

@@

For a well-designed spectrograph and sufficiently accurate model spectrum, this
{\it global} (i.e., present throughout the spectrum) covariant structure should
have a relatively low amplitude and small correlation length.  To describe that
structure, we assume a stationary covariance kernel (also called a radial basis
function) with an amplitude that depends only on the velocity distance between
two pixels, 
\begin{equation}
  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
\end{equation}
where $c$ is the speed of light.  The kernel is chosen to parametrically describe the covariance 
between pixel residuals, such that
\begin{equation}
  \Kglobal(\lambda_i, \lambda_j) =  \langle \vR_i \; \vR_j \rangle.
  \label{eqn:expectation}
\end{equation}
A variety of kernels have been used in the field of Gaussian
processes to parameterize such a mildly covariant structure
\citep[e.g.,][]{rasmussen05}.  Here we adopt the \matern\ kernel with 
$\nu = 3/2$,
\begin{equation}
  \Kglobal(\lambda_i, \lambda_j \, |\, a_{\rm g},\, \ell) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
   \label{eqn:global}
\end{equation}
which is parameterized by an amplitude $a_g$ and scale $\ell$ and makes for a smooth transition to 
negligible covariance at large $r$.  To ensure that $\vC$ remains a relatively sparse matrix that 
enables computational expendiency, we employ a Hann window function
\begin{equation}
  w\,(\lambda_i, \lambda_j |\, r_0) = \left \{ 
    \begin{array}{cc}
    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
    0 & r_{ij} > r_0 \\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
to taper the kernel (so the effective kernel is $w \, \Kglobal$).  The truncation distance 
$r_0$ can be fixed to some reasonable multiple of the scale parameter (here we set $r_0 = 4\ell$).  
The examples in Figure \ref{fig:matern} demonstrate that such a kernel readily produces correlated 
structure in the residual spectrum that is similar to those seen in a typical data--model 
comparison.

@@

Assuming that these local residuals are produced primarily by pathological differences in the 
spectral line strength (rather than shape or center), a simple Gaussian is a reasonable residual 
model.  In that case, the $k^{\rm th}$ such local residual can be described as
\begin{equation}
\vR_{\lambda}(a_k, \mu_k, \sigma_k) = \frac{a_k}{\sqrt{2\pi}{\sigma_k}} \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
\end{equation}
with an amplitude $a_k$, a mean wavelength $\mu_k$, and a width $\sigma_k$.  Following 
Eq.~\ref{eqn:expectation}, the kernel that describes the covariance of any two pixels related to 
the $k^{\rm th}$ residual region is
\begin{multline}
  \Klocal(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k) = \\
  \frac{1}{2\pi} \left(\frac{a_k}{\sigma_k}\right)^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ].
\label{eqn:klocal}
\end{multline}
We again taper this kernel with a Hann window (Eq.~\ref{eqn:Hann}) to ensure computational 
efficiency with a sparse covariance matrix; in this case, the truncation distance $r_0$ can be set 
to some multiple of the Gaussian width (we set $r_0 = 4\sigma_k$).  
Figure \ref{fig:region} demonstrates how this non-stationary Gaussian 
kernel generates a localized region of enhanced variance that successfully mimics the kind of 
residuals produced by an inaccurate spectral line model.  In effect, these kernels down-weight the 
influence of such strong residuals in the likelihood calculation, mitigating any potential bias 
they might induce on inferences of the interesting parameters ($\stellarParameters$).  In essence, this is a 
robust, flexible, and unbiased method for (correlated) outlier rejection that preserves the 
integrity of the probabilistic framework being developed (as opposed to manual or threshold-based 
clipping/masking).

These local kernels can be further modified to account for more complex residual structures.  For 
example, late-type stars with imperfectly modeled molecular bandheads may produce a complicated 
pattern of positive and negative residuals or a pronounced mismatch over a relatively large 
spectral scale.  This phenomenologically different local covariance behavior can still be treated 
in this framework, if we design an appropriate covariance kernel. Given the behavior of molecular bandheads, a \matern\ kernel tapered by the Gaussian kernel in Eq.~\ref{eqn:klocal} works well. 
\todo{We'll need to round out the Gl 51 example in Sect 3.2 before this text is finished.}

@@

@@

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[draft, width=3in, height=3in]{figs/global_posteriors.pdf}
  \includegraphics[draft, width=3in, height=3in]{figs/line_posteriors.pdf}
  \caption{\textbf{Left}: The posterior probability function of the global covariance parameters, as explored by the MCMC Gibbs sampler, marginalized over the stellar and line noise parameters. \textbf{Right}: The posterior probability function of the parameters for a single line kernel, marginalized over the stellar and global noise parameters.}
\label{fig:noise_posterior}
\end{center}
\end{figure*}

>>>>

%\vt stands for ``vector theta''
\newcommand{\vt}{ {\bm \theta}}
\newcommand{\vT}{ {\bm \Theta}}
\newcommand{\vg}{\vt_{\star, {\rm grid}}}
\newcommand{\vpp}{\vt_{\star, {\rm post}}}
\newcommand{\finst}{f_{\lambda, {\rm inst}}}
\newcommand{\fsynth}{f_{\lambda, {\rm synth}}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vtstar}{\vt_{\star}}
\newcommand{\vtcheb}{\vt_{\rm Cheb}}
\newcommand{\vtglobal}{\vt_{\rm global}}
\newcommand{\vtorder}[1]{\vt_{{\rm order}_{#1}}}
\newcommand{\vtorders}{\vt_{\rm orders}}
\newcommand{\vtline}[1]{\vt_{{\rm line}_{#1}}}
\newcommand{\vtlines}{\vt_{\rm lines}}
\newcommand{\vtcov}{\vt_{\rm cov}}
\newcommand{\vM}{\mathsf{M}}
\newcommand{\vD}{\mathsf{D}}
\newcommand{\vR}{\mathsf{R}}
\newcommand{\vC}{\mathsf{C}}
\newcommand{\fM}{ \vec{{\bm M}}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ \vec{{\bm D}}}
\newcommand{\fDi}{D_i}
\newcommand{\fR}{ {\bm R}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}
\newcommand{\matern}{Mat\'{e}rn}

@@

There are various approaches to synthesizing a spectrum, $f_{\lambda}$, for a specific set of model
parameters, $\vt_{\ast}$.  In an ideal case, a model stellar atmosphere is constructed and then 
subsequently processed through a radiative transfer code \citep[e.g.,][]{kurucz93,hauschildt99}.  
However, in general this approach is still computationally prohibitive for any iterative method of 
probabilistic inference.  One partial compromise is to interpolate over a library of atmosphere 
structures that were pre-computed for a discrete grid of parameter values, $\gridParameters$, for some arbitrary $\fundamentalParameters$, and then perform a radiative transfer calculation with 
that interpolated atmosphere to synthesize $\flam$ \citep[e.g., as for {\tt 
SME};][]{valenti96}.  A more common variant is to instead rely on interpolation over a library of 
pre-synthesized model spectra, $\flam(\gridParameters)$ \citep[e.g.,][]{castelli04,
allard12,husser13}.  While technically the former approach is most similar to the ideal case, the 
computational cost of repeated spectral synthesis is sufficiently high to make a detailed 
exploration of parameter space (particularly for data with a large spectral range) considerably 
less appealing.  A related, but different approach is to eschew forward modeling entirely (and 
therefore repeated spectral syntheses and/or library interpolations), and instead evaluate the 
models only at the discrete grid points of the library.  Then, these discretized samples of the 
posterior probability density can be interpolated to an arbitrary $\fundamentalParameters$ to construct 
appropriate confidence intervals \citep[similar to the method of {\tt SPC};][]{buchhave12}.  The 
difficulty with this latter approach is that the parameter uncertainties can be smaller than the 
grid spacing; in that case, there is valid concern that this interpolation might not accurately 
recover intrinsic parameter degeneracies.

Here we opt to take the computationally expedient approach that employs a library of model spectra, 
$\flam(\gridParameters)$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
(in practice, the metallicity $Z$ is often parameterized by [Fe/H]).  However, it is worth noting 
that the techniques we will develop are applicable to {\it any} reasonable ``front-end" that 
generates a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
$\vt_{\ast}$ must be interpolated from among the spectral library, 
\begin{equation}
\flam(\gridParameters) \leadsto \flam(\fundamentalParameters),
\label{eqn:interp}
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesireable level of inaccuracy (particularly in the $Z$ dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
calculated location in $\{\vt_{\ast}\}^{\rm grid}$, and then comparing the interpolated spectrum 
with the corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
be accurate within a few percent per high resolution model pixel.  Ideally, this pre-interpolation 
could be avoided if the spectral library was computed over a refined grid (with a substantial 
up-front computational investment); but for the time being, we can empirically propagate these 
interpolation uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).  
\comm{I see nothing was done here regarding your updated approach; how should we handle that?}

@@

We can further divide $\observationalParameters$ into those parameters that impact the model primarily in the 
spectral or flux dimensions; $\observationalParameters = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
former, we consider three kernels that contribute to the observed line-of-sight velocity 
distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, treats the instrumental 
spectral broadening.  For illustrative purposes we assume $\mathcal{F}_v^{\rm inst}$ is a Gaussian 
with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm rot}$, characterizes the 
broadening induced by (projected) stellar rotation, parameterized by $v\sin{i}$ as described by 
\citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm dop} = \delta(v-v_z)$, 
incorporates the radial velocity through a Doppler shift.  The model spectrum is modified by the 
parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
a convolution in velocity-space,\footnote{In practice, these convolutions are performed as 
multiplications in Fourier-space to better preserve spectral information \citep[cf.,][]{tonry79}; 
the mathematical formalism is presented here for clarity.}
\begin{equation}
\flam(\fundamentalParameters, \vt_{{\rm obs}, v}) = \flam(\fundamentalParameters) \otimes \varphi_v = \flam(\fundamentalParameters) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
\label{eqn:broadening}
\end{equation}
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation}
\flam(\fundamentalParameters, \vt_{{\rm obs}, v}) \mapsto \vM(\fundamentalParameters, \vt_{{\rm obs}, v}),
\label{eqn:resampling}
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-dimensional array $\vM$ (and $N_{\rm pix}$ is the number of pixels in the spectrum).  
For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
post-processing steps.

\begin{figure}[!t]
\begin{center}
  \includegraphics{figs/kernels.pdf}
  \includegraphics{figs/high2low.pdf}
  \caption{Panel 1: the instrumental (${\rm FWHM} =6.8\;\kms$) and rotational ($v \sin i = 5\;\kms$) broadening kernels, and their composite. Panel 2: A section of the raw, high resolution synthetic model at $R=$500,000. Panel 3: The spectrum after applying the composite broadening kernel, $\varphi_v$. Panel 4: The broadened spectrum downsampled to the pixels of the detector.}
\label{fig:broadening}
\end{center}
\end{figure}

At this stage, the model is further modified in the flux dimension.  A typical synthetic spectrum 
is computed as the flux that would be measured {\it at the stellar surface}, and so needs to be 
diluted by the subtended solid angle, $\Omega = (R_{\ast}/d)^2$, where $R_{\ast}$ is the stellar 
radius and $d$ is the distance.  An additional wavelength-dependent scaling factor is applied to 
account for interstellar extinction, assuming some previously-derived extinction law $A_{\lambda}$ 
\citep[e.g.,][]{cardelli89} that is parameterized by $A_V$.  The parameters $\vt_{{\rm obs}, f} = 
[\Omega, \,\, A_V]$ are applied as
\begin{equation}
\vM(\vt) \equiv \vM(\vt_{\ast}, \vt_{\rm obs}) = \vM(\vt_{\ast}, \vt_{{\rm obs}, v}) \times \Omega \times 10^{-0.4\,A_{\lambda}},
\label{eqn:scaling}
\end{equation}
where we have simplified the notation by writing $\vt \equiv [\vt_{\ast}, \,\, \vt_{\rm obs}]$.

@@

We adopt a somewhat analogous approach to deal with this issue, but cast it {\it internal} to the 
modeling framework to appropriately propagate the uncertainty introduced by additional degrees of 
freedom in the model.  The residual calibration errors are treated as an explicit contribution to 
the model spectrum, enabling us to explore the distribution of possible calibrations (see 
Section~\ref{subsec:MCMC}) and then eventually marginalize out the associated nuisance parameters.  
In essence, the inferences for the relevant stellar parameters will properly account for the 
underlying uncertainty in the calibration process.  For a continuous spectrum,
such as a long-slit spectrum or a single spectral order of an echelle spectrum,
this is achieved in practice by distorting the model spectrum with a
(low-degree) Chebyshev polynomial \citep[e.g.,][]{eisenstein06,koleva09}, 
\begin{equation}
\vM(\stellarParameters, \cheb) = \vM(\stellarParameters) \times \sum_n c_n \, T_n,
\label{eqn:chebyshev}
\end{equation}
where $T_n$ are the standard Chebyshev functions of degree $n$ and the coefficients are treated as a
set of nuisance (hyper-)parameters, $\cheb = \{c_0, c_1, \ldots, c_{n-1} \}$. For an echelle spectrum with $N$ spectral orders, there is a Chebsyhev polynomial for each order of the spectrum $i$, such that the total collection of Chebyshev parameters is $\Cheb = \{\chebi{1}, \chebi{2}, \ldots, \chebi{N} \}$.  With judicious priors on $\Cheb$, we can ensure that the 
unintended treatment of real spectral features (e.g., broad, deep molecular bands) as calibration 
artifacts is negligible (see Section~\ref{sec:examples} for examples).  The lowest-degree (scaling) 
coefficient, $c_0$, is by its nature degenerate with the solid angle parameter, $\Omega$.  
Therefore, we enforce an additional constraint by requiring that the mean of the polynomial is 
unity.  For data with a single spectral order, this means setting $c_0 = 1$.  If the goal is to 
model multiple spectral orders, we assign $c_0 = 1$ in an arbitrary order as an anchor, but permit 
the $c_0$ values in other orders to be different (as necessary).  It is worth noting that this 
formalism can, in principle, be extended to develop models for completely uncalibrated spectra, 
rather than the residual calibration mismatches as described here (with a suitable relaxation of 
the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:chebyshev} offers a practical demonstration of 
how these nuisance parameters are applied.

@@

The quality of the model spectrum is assessed by comparing to the data with a pixel-by-pixel 
likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding residual 
spectrum (of $N_{\rm pix}$ pixels) can be defined for any input parameter set,
\begin{equation}
\vR \equiv \vR(\stellarParameters, \cheb) \equiv \vD-\vM(\stellarParameters, \cheb).
\end{equation}
To quantify the probability of the data conditioned on the model, we adopt a standard 
multi-dimensional Gaussian likelihood function,
\begin{equation}
p(\vD|\vM) =  \frac{1}{[(2 \pi)^{N_{\rm pix}} \det(\vC)]^{1/2}} \exp\left ( -\frac{1}{2} \,
   \vR^\trans \vC^{-1} \vR \right ), 
   \label{eqn:likelihood}
\end{equation}
which penalizes models that yield larger residuals and explicitly allows for covariances in the 
residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix, $\vC$.  For practical 
numerical reasons, we generally consider the log-likelihood as the fit quality metric, which in 
this case is
\begin{equation}
  \ln{p(\vD | \vM)} = -\frac{1}{2} \left( \vR^\trans \vC^{-1} \vR + \ln{\det{\vC}} + N_{\rm pix} \ln{2\pi} \right).
  \label{eqn:lnlikelihood}
\end{equation}

@@

For a well-designed spectrograph and sufficiently accurate model, this {\it global} (i.e., present 
throughout the spectrum) covariant structure should have a relatively low amplitude and small 
correlation length.  To describe that structure, we use a stationary covariance kernel (or radial 
basis function) with an amplitude that depends only on the velocity distance between two pixels, 
\begin{equation}
  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
\end{equation}
where $c$ is the speed of light.  The kernel is chosen to parametrically describe the covariance 
between pixel residuals, such that
\begin{equation}
  \Kglobal(\lambda_i, \lambda_j) =  \langle \vR_i \; \vR_j \rangle.
  \label{eqn:expectation}
\end{equation}
A variety of kernels have been used in the field of Gaussian processes to parameterize such a 
mildly covariant structure \citep[e.g.,][]{rasmussen05}.  We adopt the \matern\ kernel with $\nu = 
3/2$,
\begin{equation}
  \Kglobal(\lambda_i, \lambda_j \, |\, a_{\rm g},\, \ell) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
   \label{eqn:global}
\end{equation}
which is parameterized by an amplitude $a_g$ and scale $\ell$ and makes for a smooth transition to 
negligible covariance at large $r$.  To ensure that $\vC$ remains a relatively sparse matrix that 
enables computational expendiency, we employ a Hann window function
\begin{equation}
  w\,(\lambda_i, \lambda_j |\, r_0) = \left \{ 
    \begin{array}{cc}
    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
    0 & r_{ij} > r_0 \\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
to taper the kernel (so the effective kernel is $w \, \Kglobal$).  The truncation distance 
$r_0$ can be fixed to some reasonable multiple of the scale parameter (here we set $r_0 = 4\ell$).  
The examples in Figure \ref{fig:matern} demonstrate that such a kernel readily produces correlated 
structure in the residual spectrum that is similar to those seen in a typical data--model 
comparison.

@@

Assuming that these local residuals are produced primarily by pathological differences in the 
spectral line strength (rather than shape or center), a simple Gaussian is a reasonable residual 
model.  In that case, the $k^{\rm th}$ such local residual can be described as
\begin{equation}
\vR_{\lambda}(a_k, \mu_k, \sigma_k) = \frac{a_k}{\sqrt{2\pi}{\sigma_k}} \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
\end{equation}
with an amplitude $a_k$, a mean wavelength $\mu_k$, and a width $\sigma_k$.  Following 
Eq.~\ref{eqn:expectation}, the kernel that describes the covariance of any two pixels related to 
the $k^{\rm th}$ residual region is
\begin{multline}
  \Klocal(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k) = \\
  \frac{1}{2\pi} \left(\frac{a_k}{\sigma_k}\right)^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ].
\label{eqn:klocal}
\end{multline}
We again taper this kernel with a Hann window (Eq.~\ref{eqn:Hann}) to ensure computational 
efficiency with a sparse covariance matrix; in this case, the truncation distance $r_0$ can be set 
to some multiple of the Gaussian width (we set $r_0 = 4\sigma_k$).  Figure \ref{fig:region} 
demonstrates how this non-stationary Gaussian kernel generates a localized region of enhanced 
variance that successfully mimics the kind of residuals produced by an inaccurate spectral line 
model.  These kernels down-weight the influence of such strong residuals in the likelihood 
calculation, mitigating any potential bias they might induce on inferences of the interesting 
parameters ($\stellarParameters$).  In essence, this is a robust, flexible, and unbiased method for (correlated) 
outlier rejection that preserves the integrity of the probabilistic framework being developed (as 
opposed to manual or threshold-based clipping or masking).

\comm{These local kernels can be further modified to account for more complex residual structures.  For 
example, late-type stars with imperfectly modeled molecular bandheads may produce a complicated 
pattern of positive and negative residuals or a pronounced mismatch over a relatively large 
spectral scale.  This phenomenologically different local covariance behavior can still be treated 
in this framework, if we design an appropriate covariance kernel. Given the behavior of molecular bandheads, a \matern\ kernel tapered by the Gaussian kernel in Eq.~\ref{eqn:klocal} works well.}
\todo{We'll need to round out the Gl 51 example in Sect 3.2 before this text is finished.}

@@

\subsection{Exploring the Posterior} \label{subsec:MCMC}

@@

To explore the posterior probability distribution and make inferences on the parameters of 
interest, we employ a hierarchical Monte Carlo Markov Chain (MCMC) approach that nests the standard 
Metropolis-Hastings algorithm (for sampling the parameters $\vt$) inside the Gibbs sampler (which 
rotates through subsets of the hyper-parameters $\vT$).
